### 2021-10-24 plos_latex_template.tex
Federico's edits

The authors of this article participated in the organization of the annual user conference of the R Project for Statistical Computing, held in July 2021. useR! conferences are non-profit events organized by volunteers from the R community and managed by the R Foundation. The conference attracts a broad range of participants from academia, industry, government, and the non-profit sector. For 2021, we aimed to build a high-quality virtual and explicitly global conference in a kind, inclusive, accessible, and welcoming environment for everyone. 
In this article, we streamline our most important learnings in 10 simple rules to host an inclusive conference. These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level. The rules were learned during a virtual conference, but serve also as guidelines to in-person and hybrid events. \fede{A comment from Fede} \rocio{an ideal reply to that from Rocio} \doro{Doro chiming in, before the comment is resolved}

@@

% Rocío: Main message of the paragraph: Why we are writing this piece
Conferences, from the Latin \textit{conferre}--`to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields. 
Lack of representation and unwelcoming--or overtly aggressive--environments exclude people from the opportunities for learning, sharing, and producing knowledge, and becoming active community members, all of which are readily provided to non-marginalized people \cite{hendersonThoughtfulGatheringsGendering2020}.
Exclusionary conference experiences can divert career paths, affect lives, and drive people out of academia \cite{biggsAcademicConferenceChilly2018, hendersonThoughtfulGatheringsGendering2020}. 
This message is not new. 
Barriers such as unattainable registration costs, sexism, and ableism, among others, have already been exposed and discussed in the literature \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}, 
and some proposals for more inclusive conferences have been put into practice \cite{gichoraTenSimpleRules2010a, levitisCenteringInclusivityDesign2021, atkinsonJournalMedicine20202021, foramittiVirtuesVirtualConferences2021, ninerBetterWhomLeveling2021, rabyMovingAcademicConferences2021}, with a primary focus on the online format to open doors for inclusion. %https://www.cscce.org/resources/organizing-community-events/

% Rocío: Main message of the paragraph: Roadmap of the paper
This article puts the focus on practices for inclusive conferences, whether they are virtual or in-person, though we advocate for having a strong online component, well-articulated with the in-person component, for all conferences.
The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
The rules can also be helpful to local/virtual organizers who desire to make an inclusive conference starting at the planning stage.
These tips stem from the authors' experience of organizing useR! 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}. 
We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many persons as possible. 
Here, we share the lessons learned within the past year of organizing useR! 2021, summarized as ten simple rules towards hosting an inclusive conference.
The rules are organized in three groups (Figure \ref{fig:diagram}). %original source: https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing
%sara: still need to better format figure size and get the alt text right :P
Group 1 includes rules 1, 2, and 10, which refer to pillars of an inclusive conference: embracing diversity in all its dimensions, creating a safe and welcoming environment for everyone, and making the conference part of a long-term process for inclusion.
The second group includes rules 3 and 4 and focuses on the people who participate in the conference. 
Rule 3 refers to the importance of working with an inclusive and diverse organizing team, and Rule 4 concerns the necessity of removing implicit and systemic bias from spotlight roles like keynote speakers, other presenters, program committee members, or other session chairs. 
The third group includes rules 5 to 9. These rules are about components of the conference that should be carefully planned for: an online component, accessibility to people with disabilities, language inclusiveness, a welcoming communication strategy, and financial resources to support inclusion. 
These rules apply equally to academic, industry, or mixed conferences. While these rules are inspired in a global virtual experience they also apply at the regional or local level and can inform in-person and hybrid events.

@@

% Rocío: Main idea: what is diversity and why does it matter? (inequalities)
Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
Human diversity should be celebrated and respected in every way. Nonetheless, we live in a world with implicit hierarchies along these axes. Some statuses (e.g., cisgender, white, male, from the United States or Western Europe) hold the privilege of being the defaults centered around which all systems--including conferences--are consciously and unconsciously built.  
While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.

% Rocío: more effort towards including the more excluded
Recognizing our privileges--unearned advantages given by society to some people but not all--particularly in our field and in our scientific or professional community, will help identify which subgroups have been the most excluded or discriminated against. 
These are the groups we need to make more effort to include. 
Investing more effort in the most excluded groups does not mean neglecting the others, but it will guide the vision of diversity for your conference--and your strategies to achieve it.
Will a more diverse conference translate into a gender distribution of your speakers that is representative of the general population? 
Would it be the presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees? 
Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions?
%"The answer to these questions depend on your field, region or community.
Intersectionality, or the complex, cumulative way in which the effects of multiple forms of discrimination combine or intersect, should be taken into account.

@@

% Rocío: main idea: Code of conduct for a safe place
Adopting a code of conduct and gathering a team to enforce it are key aspects in creating a safe environment during a conference \cite{favaroYourScienceConference2016}.
%Dorothea: it is not a document, right? The document will not keep the community safe. Maybe say "a set of rules" in case someone doesn't know what a code of conduct is?
The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behavior, and the way to report violations \cite{auroraHowRespondCode2019}. 
The code of conduct team must receive training on how to receive reports, respond to incidents, and communicate their responses, and organize accordingly. 
A diverse code of conduct team will be more understanding of power dynamics and sensitive to discrimination and harassment issues. 
There are also disciplinary cultures and geographical considerations that may need to be considered when developing and communicating the code of conduct. 
We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for the code of conduct teamwork.
% Rocío: two of these sentences, diverse CoC team, and communication of CoC, may belong to rules organizing team and communication, respectively.

@@

% Rocío: Idea of this paragraph: build a diverse team
A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team, in charge of the conference decision-making.
Build a team with people from different regions, genders, races and ethnicities, socioeconomic statuses, and other aspects of diversity. 
Go beyond balancing all genders in this effort and pay special attention to to intersectional diversity (see \textbf{Rule \ref{rule_diversity}}). 
To ensure a deep understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the general population.
If you already have an organizing team, check for gaps in its composition.

@@

% Rocío: Idea of this paragraph: positive outcomes of having a diverse team - keeping it general at least
Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams--when you are not used to them--but the positive outcomes are worthwhile. 
Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}. 
In addition, a diverse team plays an important role at creating a welcoming space because representation--seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes--is one of the best ways to create a sense of belonging for everyone participating in the conference (see \textbf{Rule \ref{rule_inclusion}}).

@@

% Rocío: main idea: unbias roles and give spotlight to the roles--and people who have contributed in roles--that are not stereotypically categorized as success
In academia and the industry, not only is the list of people considered in spotlight roles usually biased, but the spotlight roles themselves are.
Scientific publication and development of software, are often regarded with higher value than community building or teaching, which may be equally if not more challenging, less rewarded, and usually tackled by women, people of color, people with disabilities and from other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020}.
Defy the stereotypical criteria for success by acknowledging these community practices and the people behind them.
You can, for instance, give more space in your conference to talks about community building as an area of knowledge production, and reframe \doro{amend?} the awards ceremony to acknowledge those who prepared accessible slides and presentations (see Rule \ref{rule_accessibility}), for being mindful of inclusiveness. \doro{Suggestion: "amend the awards ceremony to also acknowledge those who prepared accessible slides and presentations, contributed to community building or presented teaching materials"}
Finally, do not restrict people from marginalized groups or community-builders to talk or work only on issues related to diversity and inclusion; recognize their areas of expertise and interest and offer them space to showcase them.

@@

% Rocío: main idea of the paragraph: online conferences can open opportunities for inclusion
In-person interaction in conferences is priceless, but only for the ones who can afford to attend. 
Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019,jooKeepOnlineOption2021}. 
Online conferences are more inclusive because they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities \cite{salibaGettingGripsOnline2020}.
This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020, jooKeepOnlineOption2021}.
The online format may also make it easier to be inclusive of geographic regions by encompassing several timezones in the week or easily deciding to rotate the favored timezone year to year without depending on a conference central location.

% Rocío: main idea of the paragraph: In the context of reopening, hybrid conferences may be a thing. They also need to be inclusive.
Within the context of the COVID-19 pandemic, many conferences embraced the online format; but at the time of writing, some are reverting to in-person, which risks going back to the barriers mentioned above.
Alternatively, a conference could have a hybrid format with an in-person and an online components that articulate well into a hybrid format that maximizes the experience of everyone attending your conference. This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely. 
The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.
%Needless to say, an online component is not inclusive \textit{per se} without taking many other precautions, and it does not eliminate the need to organize an inclusive in-person component. 
%For instance, to avoid favoring the same regions in every edition, we suggest rotating the location of the conference—if it is global, moving from one continent to another. 
% Rocío: Now I'm not sure if we need the last two sentences that I commented.

% Rocío: main idea: Networking may seem the weakest part of online conference, but it doesn't have to be. 
% Rocío: I'm still doubting if this should be the 2nd paragraph, and I think that it needs some editing.
Networking and socializing have been mentioned as challenging aspects of online conferences, mostly because we are used to interactions at in-person social settings such as coffee or lunch breaks \cite{salibaGettingGripsOnline2020, roosOnlineConferencesNew2020}. 
However, in-person settings might not be comfortable or appealing to everyone for socializing; 
some of these spaces may be exclusionary (e.g. galas or dinner nights at expensive venues), and common place for code of conduct violations to occur \cite{auroraHowRespondCode2019}. 
On the other hand, there is evidence that virtual communication before face-to-face interaction can make people from minoritized groups feel more included, thus participate more (e.g. \cite{trianaDoesOrderFacetoFace2012, blackEngenderingBelongingThoughtful2020}). 
Organizers of online conferences should invest time in creating opportunities to meet and bond virtually, respecting people's limits, preferences, and remembering that `the usual' does not necessarily work for everyone, and that one single networking activity will never serve the whole community. 
It is worth trying new and varied activities that might work with subgroups of attendees.
Some ideas in this line are: offering the option of written chat only, instead of voice or video conversations, opening events with teamwork like trivia, offer some events that can be enjoyed passively like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session.

@@

Conferences are among the least accessible spaces that people with disabilities may encounter \cite{priceAccessImaginedConstruction2009}. Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}. This is one key aspect where having someone in the organizing team allows them to take part in the decisions from the beginning (see \textbf{Rule \ref{rule_organizing_team}}), as thinking about disability or simulating it are not substitutes for real-life experience \cite{costanzachockDesign2020}. Planning for accessibility requires time and early decision-making \cite{irishIncreasingParticipationUsing2020}. As other aspects of inclusion, dealing with accessibility at the last minute is the recipe for a disastrous conference. If you do not consider accessibility from the conference inception, it is highly likely you will be better off without trying to patch accessibility at the last minute. Key decisions in this respect are hard to correct. 
For in-person conferences, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live interpretation, just to name a few important features. In addition to this, the organizers should take care proactively of invisible disabilities, for example, by providing quiet spaces for privacy and noise-free conversations, provide chairs in open spaces, and menus that include sugar-free, gluten-free, and alcohol-free options.

Regardless of the conference format, all platforms (website, chat, conference administration tools) and images used for the communication strategy of the conference should be screen-reader friendly and keyboard accessible, and have alternative text. Any videos should have captions and a transcript. The organizing team should provide accessibility guidelines for slides and presentations, encourage their use, and be available for any questions presenters and attendees may have. Slide decks should be made available beforehand, either in webpages or available for download to ensure that everybody can follow their content during the presentations.

Accessibility practices must include social events and networking, and include activities that do not restrict participation based on body type or ability. Importantly, all these accessibility practices are inclusive not only for people with disabilities but to everyone %ast: refs here.
For instance, captions are helpful for non-native speakers, having the material available for download helps attendees with low bandwidth connection, etc. Accessibility efforts should be explicitly displayed in the official communications of the conference (see \textbf{Rule \ref{rule_communication}}).

@@

% Rocío: main idea: English as the only language makes some people privileged and is a barrier for others
In international conferences, the linguistic diversity of conference participants is often overlooked. 
English is usually the official and sole language for submissions, presentations, tutorials, workshops, conference platforms, webpage, and official communications. 
While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, this makes being a native English speaker, a privilege.
Non-native English speakers could miss opportunities to attend or actively participated in conferences (e.g. ask questions or participated in discussions),
and conferences may miss innovative contributions.

@@

%Rocío: Main idea: Include people in your communication 
Your communication strategy should reflect the spirit of the conference and be inclusive by design. 
As you design the promotion plan, actively reach out and promote the conference to people who have been systematically excluded. 
If part of the community has been historically discriminated against, emphasize that they are particularly welcome in this event and that the organization will make it a safe and inclusive environment. This could be done through publishing the diversity statement (e.g., https://user2021.r-project.org/about/diversitystatement/), the code of conduct, and accessibility guidelines and advertising financial support, or other creative ways that show that everyone is seen, respected, and welcome (Fig. MARMOTS).
Also, try to advertise the conference in multiple languages (\textbf{Rule \ref{rule_language}}) and multiple platforms (e.g. Twitter, Facebook, LinkedIn, conference website, email lists)--
part of your community may use a different platform than the one who regularly attend the conference. 
With proactive inclusive communication strategies, potential attendees will be assured that this is their space and community too (e.g. having accessibility guidelines on your website sends a welcoming message to people with disabilities), without the emotional labor of contacting the organizers and be public about your commitment to equity and inclusion.

@@

\section{Allocate financial resources to support your conference goals}
\label{rule_financial}

% Rocío: main idea: Budgets are limited and need to choose priorities.
Conference budgets are limited and rely mostly on sponsorship and attendance fees. 
While some expenses might be more or less fixed, % Rocío: can we give examples?
% Dorothea: this depends so much on the conference type. I wanted to write "venue"but this is not true of course, so I don't know what to add. 
allocation of resources has to be intentional to support the goals of the conference towards inclusion. 
It is important to estimate the costs for these inclusive practices and define your priorities in advance, like paying the organizing team, necessary training (e.g. code of conduct), and accessibility practices (e.g. captions). 
Aditional support for attendees should also be considered in the budget: child care support, transportation fees, visa-related support (if in-person), internet connection services (for the virtual component). 
Consider that an online conference might reduce organization costs (e.g. no rental costs for a physical venue), allowing to redirect the money towards other inclusive priorities. 
Instead of depending on registration fees, ask sponsors to support specific items in your inclusion expenses.
%Dorothea: I slightly disagree with that one. Maybe just leave the first part out?

>>>>

The authors of this article participated in the organization of the annual user conference of the R Project for Statistical Computing, held in July 2021. useR! conferences are non-profit events organized by volunteers from the R community and managed by the R Foundation. The conference attracts a broad range of participants from academia, industry, government, and the non-profit sector. For 2021, we aimed to build a high-quality virtual and explicitly global conference in a kind, inclusive, accessible, and welcoming environment for everyone. 
In this article, we streamline\fede{outline} our most important learnings\fede{lessons learnt} in 10 simple rules to host an inclusive conference.\fede{general comment on this: we should clearly state that this whole-virtual setting was a clear consequence of the pandemics, AND it was planned like this from the beginning, and not changing in-between to this format} These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level. The rules were learned during a virtual conference, but serve also as guidelines to in-person and hybrid events. \fede{A comment from Fede} \rocio{an ideal reply to that from Rocio} \doro{Doro chiming in, before the comment is resolved}

@@

% Rocío: Main message of the paragraph: Why we are writing this piece
Conferences, from the Latin \textit{conferre}--`to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.\fede{between this and the next sentence, there is a strong break. need some ligation, something like "Despite the many efforts in this direction, ..."}
However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields. 
Lack of representation and unwelcoming--or overtly aggressive--environments exclude people from the opportunities for learning, sharing, and producing knowledge, and becoming active community members, all of which are readily provided to non-marginalized people \cite{hendersonThoughtfulGatheringsGendering2020}.
Exclusionary conference experiences can divert career paths, affect lives, and drive people out of academia \cite{biggsAcademicConferenceChilly2018, hendersonThoughtfulGatheringsGendering2020}. 
This message is not new. 
Barriers such as unattainable registration costs, sexism, and ableism, among others, have already been exposed and discussed in the literature \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}, 
and some proposals for more inclusive conferences have been put into practice \cite{gichoraTenSimpleRules2010a, levitisCenteringInclusivityDesign2021, atkinsonJournalMedicine20202021, foramittiVirtuesVirtualConferences2021, ninerBetterWhomLeveling2021, rabyMovingAcademicConferences2021}, with a primary focus on the online format to open doors for inclusion.\fede{do we need maybe a short definition of the concepts of diversity-equity-inclusion for us/refer to something?} %https://www.cscce.org/resources/organizing-community-events/

% Rocío: Main message of the paragraph: Roadmap of the paper
This article puts the focus on practices for inclusive conferences, whether they are virtual or in-person, though we advocate for having a strong online component, well-articulated with the in-person component, for all conferences.\fede{instead of "whether they are" -> ", which can be applied to both commonly existing formats, ..."; the fact we suggest to have an online component is not so tight to this, I guess?}
The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
The rules can also be helpful to local/virtual organizers who desire to make an inclusive conference starting at the planning stage.
These tips stem from the authors' experience of organizing useR! 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}. 
We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many persons as possible. 
Here, we share the lessons learned within the past year of organizing useR! 2021, summarized as ten simple rules towards hosting an inclusive conference.
The rules are organized in three groups (Figure \ref{fig:diagram}). %original source: https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing
%sara: still need to better format figure size and get the alt text right :P
Group 1 includes rules 1, 2, and 10, which refer to pillars of an inclusive conference: embracing diversity in all its dimensions, creating a safe and welcoming environment for everyone, and making the conference part of a long-term process for inclusion.
The second group includes rules 3 and 4 and focuses on the people who participate in the conference. 
Rule 3 refers to the importance of working with an inclusive and diverse organizing team, and Rule 4 concerns the necessity of removing implicit and systemic bias from spotlight roles like keynote speakers, other presenters, program committee members, or other session chairs. 
The third group includes rules 5 to 9. These rules are about components of the conference that should be carefully planned for: an online component, accessibility to people with disabilities, language inclusiveness, a welcoming communication strategy, and financial resources to support inclusion. 
These rules apply equally to academic, industry, or mixed conferences. While these rules are inspired in a global virtual experience they also apply at the regional or local level and can inform in-person and hybrid events.
\fede{maybe something like: "Even if this set of rules is derived from our experience with a global scale event, we believe these can be easily transferred to events with regional or local scopes"}

@@

% Rocío: Main idea: what is diversity and why does it matter? (inequalities)
Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few\fede{we could substantiate this with one or two refs?}.
Human diversity should be celebrated and respected in every way. Nonetheless, we live in a world with implicit hierarchies along these axes. Some statuses (e.g., cisgender, white, male, from the United States or Western Europe) hold the privilege of being the defaults centered around which all systems--including conferences--are consciously and unconsciously built.  
While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.

% Rocío: more effort towards including the more excluded
Recognizing our privileges--unearned advantages given by society to some people but not all--particularly in our field and in our scientific or professional community, will help identify which subgroups have been the most excluded or discriminated against. 
These are the groups we need to make more effort to include. \fede{General thought: I like the way we say WE and YOU. I still think I saw most such 10SR papers speak to the readers addressing them with the roles. Which would be "Organizers", "Attendees". Just my 2 cents as avid reader of 10SR papers ;)}
Investing more effort in the most excluded groups does not mean neglecting the others, but it will guide the vision of diversity for your conference--and your strategies to achieve it.
Will a more diverse conference translate into a gender distribution of your speakers that is representative of the general population? 
Would it be the presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees? 
Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions?
%"The answer to these questions depend on your field, region or community.
Intersectionality, or the complex, cumulative way in which the effects of multiple forms of discrimination combine or intersect, should be taken into account.

@@

% Rocío: main idea: Code of conduct for a safe place
Adopting a code of conduct and gathering a team to enforce it are key aspects in creating a safe environment during a conference \cite{favaroYourScienceConference2016}.
%Dorothea: it is not a document, right? The document will not keep the community safe. Maybe say "a set of rules" in case someone doesn't know what a code of conduct is?
The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behavior, and the way to report violations \cite{auroraHowRespondCode2019}. 
The code of conduct team must receive training on how to receive reports, respond to incidents, communicate their responses, and organize accordingly. 
Assembling a diverse code of conduct team will be more understanding of power dynamics and sensitive to discrimination and harassment issues. 
There are also disciplinary cultures and geographical considerations that may need to be considered when developing and communicating the code of conduct. 
We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for the code of conduct teamwork.
% Rocío: two of these sentences, diverse CoC team, and communication of CoC, may belong to rules organizing team and communication, respectively.

@@

% Rocío: Idea of this paragraph: build a diverse team
A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team, in charge of the conference decision-making.
Organizers should build a team with people from different regions, genders, races and ethnicities, socioeconomic statuses, and other aspects of diversity. 
Go beyond balancing all genders in this effort and pay special attention to to intersectional diversity (see \textbf{Rule \ref{rule_diversity}}). 
To ensure a deep understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the general population.
If you already have an organizing team, check for gaps in its composition.

@@

% Rocío: Idea of this paragraph: positive outcomes of having a diverse team - keeping it general at least
Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams--especially when you are not used to them--but the positive outcomes are worthwhile. 
Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}. 
In addition, a diverse team plays an important role at creating a welcoming space because representation--seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes--is one of the best ways to create a sense of belonging for everyone participating in the conference (see \textbf{Rule \ref{rule_inclusion}}).

@@

% Rocío: main idea: unbias roles and give spotlight to the roles--and people who have contributed in roles--that are not stereotypically categorized as success
In academia and the industry, not only is the list of people considered in spotlight roles usually biased, but the spotlight roles themselves are.
Scientific publication and development of software, are often regarded with higher value than community building or teaching, which may be equally if not more challenging, less rewarded, and usually tackled by women, people of color, people with disabilities and from other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020}.
Defy the stereotypical criteria for success by acknowledging these community practices and the people behind them.
You can, for instance, give more space in your conference to talks about community building as an area of knowledge production, and reframe \doro{amend?}\fede{or, "promote a format for..."} the awards ceremony to acknowledge those who prepared accessible slides and presentations (see Rule \ref{rule_accessibility}), for being mindful of inclusiveness. \doro{Suggestion: "amend the awards ceremony to also acknowledge those who prepared accessible slides and presentations, contributed to community building or presented teaching materials"}
Finally, do not restrict people from marginalized groups or community-builders to talk or work only on issues related to diversity and inclusion; recognize their areas of expertise and interest and offer them space to showcase them.

@@

% Rocío: main idea of the paragraph: online conferences can open opportunities for inclusion
In-person interaction in conferences is priceless, but often this is true and possible only for the ones who can afford to attend. 
Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019,jooKeepOnlineOption2021}. 
Online conferences are more inclusive because they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities \cite{salibaGettingGripsOnline2020}.
This means that online conferences will have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020, jooKeepOnlineOption2021}.
The online format may also make it easier to be inclusive of geographic regions by encompassing several timezones in the week or easily deciding to rotate the favored timezone year to year, without depending on a conference central location.

% Rocío: main idea of the paragraph: In the context of reopening, hybrid conferences may be a thing. They also need to be inclusive.
Within the context of the COVID-19 pandemic, many conferences embraced the online format; but at the time of writing, some are reverting to in-person, which risks going back to the barriers mentioned above.
Alternatively, a conference could have a hybrid format with an in-person and an online components that articulate well into a hybrid format that maximizes the experience of everyone attending your conference. This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely. 
The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a "consolation prize" to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.
%Needless to say, an online component is not inclusive \textit{per se} without taking many other precautions, and it does not eliminate the need to organize an inclusive in-person component. 
%For instance, to avoid favoring the same regions in every edition, we suggest rotating the location of the conference—if it is global, moving from one continent to another. 
% Rocío: Now I'm not sure if we need the last two sentences that I commented.

% Rocío: main idea: Networking may seem the weakest part of online conference, but it doesn't have to be. 
% Rocío: I'm still doubting if this should be the 2nd paragraph, and I think that it needs some editing.
Networking and socializing have been mentioned as challenging aspects of online conferences, mostly because we are used to interactions at in-person social settings such as coffee or lunch breaks \cite{salibaGettingGripsOnline2020, roosOnlineConferencesNew2020}. 
However, in-person settings might not be comfortable or appealing to everyone for socializing; 
some of these spaces may be exclusionary (e.g. galas or dinner nights at expensive venues), and common place for code of conduct violations to occur \cite{auroraHowRespondCode2019}. 
On the other hand, there is evidence that virtual communication before face-to-face interaction can make people from minoritized groups feel more included, thus participate more (e.g. \cite{trianaDoesOrderFacetoFace2012, blackEngenderingBelongingThoughtful2020}). 
Organizers of online conferences should invest time in creating opportunities to meet and bond virtually, respecting people's limits, preferences, and remembering that `the usual' does not necessarily work for everyone, and that one single networking activity will never serve the whole community. 
It is worth trying new and varied activities that might work with subgroups of attendees.
Some ideas in this line are: offering the option of written chat only, instead of voice or video conversations, opening events with teamwork like trivia, offer some events that can be enjoyed passively like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session.\fede{another activity: virtual city tours, I had it as a corollary activity for a summer school}

@@

Conferences are among the least accessible spaces that people with disabilities may encounter \cite{priceAccessImaginedConstruction2009}. Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}. This is one key aspect where having someone in the organizing team allows them to take part in the decisions from the beginning (see \textbf{Rule \ref{rule_organizing_team}}), as thinking about disability or simulating it are not substitutes for real-life experience \cite{costanzachockDesign2020}. Planning for accessibility requires time and early decision-making \cite{irishIncreasingParticipationUsing2020}. As other aspects of inclusion, dealing with accessibility at the last minute is the recipe for a disastrous conference. If you do not consider accessibility from the conference inception, it is highly likely you will be better off without trying to patch accessibility at the last minute. Key decisions in this respect are hard to correct. 
For in-person conferences, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live interpretation, just to name a few important features. In addition to this, the organizers should take care proactively of "invisible disabilities", for example, by providing quiet spaces for privacy and noise-free conversations, provide chairs in open spaces, and menus that include sugar-free, gluten-free, and alcohol-free options.

Regardless of the conference format, all platforms (website, chat, conference administration tools) and images used for the communication strategy of the conference should be screen-reader friendly and keyboard accessible, and have alternative text. Any videos should have captions and a transcript.

The organizing team should provide accessibility guidelines for slides and presentations, encourage their use, and be available for any questions presenters and attendees may have. Slide decks should be made available beforehand, either in webpages or available for download to ensure that everybody can follow their content during the presentations.

Accessibility practices should also include social events and networking, and include activities that do not restrict participation based on body type or ability. Importantly, all these accessibility practices are inclusive not only for people with disabilities but to everyone %ast: refs here.
For instance, captions are helpful for non-native speakers, having the material available for download helps attendees with low bandwidth connection, etc. Accessibility efforts should be explicitly displayed in the official communications of the conference (see \textbf{Rule \ref{rule_communication}}).

@@

% Rocío: main idea: English as the only language makes some people privileged and is a barrier for others
In international conferences, the linguistic diversity of conference participants is often overlooked. 
English is usually the official and sole language for submissions, presentations, tutorials, workshops, conference platforms, webpage, and official communications. 
While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, this makes being a native English speaker a privilege.
Non-native English speakers could miss opportunities to attend or actively participate in conferences (e.g. asking questions or participating in discussions),
and conferences may miss innovative contributions.

@@

%Rocío: Main idea: Include people in your communication 
Your communication strategy should reflect the spirit of the conference and be inclusive by design. 
As you design the promotion plan, actively reach out and promote the conference to people who have been systematically excluded. 
If part of the community has been historically discriminated against, emphasize that they are particularly welcome in this event and that the organization will make it a safe and inclusive environment. This could be done through publishing the diversity statement (e.g., https://user2021.r-project.org/about/diversitystatement/), the code of conduct, and accessibility guidelines and advertising financial support, or other creative ways that show that everyone is seen, respected, and welcome (Fig. MARMOTS).
Also, try to advertise the conference in multiple languages (\textbf{Rule \ref{rule_language}}) and multiple platforms (e.g. Twitter, Facebook, LinkedIn, conference website, mailing lists)--
part of your community may use a different platform than the ones who regularly attend the conference. 
With proactive inclusive communication strategies, potential attendees will be assured that this is their space and community too (e.g. having accessibility guidelines on your website sends a welcoming message to people with disabilities), without the emotional labor of contacting the organizers and be public about your commitment to equity and inclusion.

@@

\section{Allocate adequate financial resources to support your conference goals}
\label{rule_financial}

% Rocío: main idea: Budgets are limited and need to choose priorities.
Conference budgets are usually limited and rely mostly on sponsorship and attendance fees. 
While some expenses might be more or less fixed, % Rocío: can we give examples?
% Dorothea: this depends so much on the conference type. I wanted to write "venue"but this is not true of course, so I don't know what to add. 
allocation of resources has to be intentional to support the goals of the conference towards inclusion. 
It is important to estimate the costs for these inclusive practices and define your priorities in advance, like paying the organizing team, necessary training (e.g. code of conduct), and accessibility practices (e.g. captions). 
Aditional support for attendees should also be considered in the budget: child care support, transportation fees, visa-related support (if in-person), internet connection services (for the virtual component). 
Consider that an online conference might reduce organization costs (e.g. no rental costs for a physical venue), allowing to redirect the money towards other inclusive priorities. 
Instead of depending on registration fees, ask sponsors to support specific items in your inclusion expenses.
%Dorothea: I slightly disagree with that one. Maybe just leave the first part out?
